The cost of undisturbed landscapes

### 2020-04-21 projects/asparagus/doc/slides.tex
updates to project asparagus slides.tex

    \section{Background \& Motivation}\label{sec:background}
    \begin{frame}
        \frametitle{State of the Art}
        Power system models used to
        \begin{itemize}
            \item demonstrate the feasibility of highly renewable energy systems
            \item determine the system cost-minimizing mix of renewable energy technologies
        \end{itemize}
        find significant capacities of renewable energy technologies (RET), largely wind power, need to be added.
        \medskip

@@

        Yet, such `binary' analysis neglects the trade-offs faced by a social planner deciding on RET expansion.
    \end{frame}

@@

        \begin{textblock*}{5cm}
            (1.0cm,2cm)
            Consider a renewable energy technology that
            \begin{itemize}
                \item saves system cost (relative to its best alternative), but
                \item comes at social cost (e.g. due to local negative external effects)
            \end{itemize}
            \medskip

            A social planner should deploy this RET (instead of another one) up to the capacity at which
            \begin{itemize}
                \item system cost saving equals
                \item added social cost
            \end{itemize}
        \end{textblock*}

        \begin{textblock*}{8.0cm}
            (6.5cm,2cm)
            \begin{tikzpicture}
                \begin{axis}[
                    width=\textwidth,
                    clip=false,
                    axis lines=left,
                    ticks=none,
                    xlabel={capacity substituted},
                    ylabel={cost},
                    domain=0.5:1.75,
                    xmin=0.5, xmax=2,
                    ymin=0, ymax=4,
                    samples=200
                ]
                    \addplot[blue, thick] (x, 1.3*x^1.1+0.6);
                    \node at (axis cs: 1.925,3) [blue, align=left] {saved\\ system\\ cost};

@@

\begin{textblock*}{10cm}(0.75cm, 1.5cm) % {block width} (coords)
\includegraphics[width=\textwidth, keepaspectratio]{./figures/undisturbed_base}
\end{textblock*}

\begin{textblock*}{4.5cm}(10.9cm, 1.5cm)
\begin{small}
\begin{table}
\begin{tabular}{p{1.8cm} | r l}
Capital cost of solar PV & $630$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}}
\end{tabular}
\end{table}
\smallskip
\quad $\rightarrow$ about $\sfrac{2}{3}$ rooftop PV, \\
\quad $\sfrac{1}{3}$ open space PV

\vspace{4.25cm}
\quad wind power and solar PV \\ \quad are substitutes
\end{small}
\end{textblock*}

\begin{textblock*}{9.5cm}(1.5cm, 8.2cm)
\begin{tikzpicture}%[thick]
    \draw (5.75,0) -- (9,0);
    \draw[{Latex[length=2mm]}-] (0,0) -- (3.25,0);
    \draw (4.5,0.0) node[] {\scriptsize more wind power};
\end{tikzpicture}

@@

\section{Sensitivity}
\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Sensitivity to capital cost of solar PV}
\begin{textblock*}{10cm}(0.75cm, 2.0cm) % {block width} (coords)
\includegraphics[width=\textwidth, keepaspectratio]{./figures/sens_nrg}
\end{textblock*}

\begin{textblock*}{4cm}(11cm, 2.25cm)
\begin{small}
    Capital cost estimates for solar PV in 2030
    \begin{table}
        \begin{tabular}{p{1.8cm} | p{0.3cm} l}
            Small-scale rooftop & $830$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
            Utility-scale open space & $280$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
        \end{tabular}
    \end{table}
\end{small}
\end{textblock*}
\end{frame}

\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Sensitivity to capital cost of solar PV}
\begin{textblock*}{10cm}(0.75cm, 2.0cm) % {block width} (coords)
\includegraphics[width=\textwidth, keepaspectratio]{./figures/undisturbed_low}
\end{textblock*}

\begin{textblock*}{4cm}(11cm, 2.25cm)
\begin{small}
    \begin{table}
        \begin{tabular}{p{1.8cm} | p{0.3cm} l}
            Capital cost of solar PV & $280$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}}
        \end{tabular}
    \end{table}
\end{small}
\end{textblock*}
\end{frame}

>>>>

    \section{Background \& Motivation}\label{sec:background}
    \begin{frame}
        \frametitle{State of the Art}
        Power system models used to
        \begin{itemize}
            \item demonstrate the feasibility of highly renewable energy systems
            \item determine the system cost-minimizing mix of renewable energy technologies
        \end{itemize}
        find significant capacities of renewable energy technologies (RET), in Central Europe largely wind power,
        need to be added.
        \medskip

@@

        Yet, this neglects the trade-offs faced by a social planner deciding on RET expansion.
    \end{frame}

@@

        \begin{textblock*}{5.75cm}
            (1.0cm,2cm)
            Consider a renewable energy technology $\omega$ that
            \begin{itemize}
                \item is \emph{system cost} minimizing, i.e.\ it saves cost relative to its best alternative $\mu$, but
                \item comes at \emph{social cost}, e.g. due to unaccounted negative external effects
            \end{itemize}
            \medskip

            A social planner should substitute $\omega$ with $\mu$ up to the point where the increase in system cost
            equals the reduction in harm (from $\omega$'s negative externality unaccounted in system cost)
        \end{textblock*}

        \begin{textblock*}{7.5cm}
            (7.5cm,2cm)
            \begin{tikzpicture}
                \begin{axis}[
                    width=\textwidth,
                    clip=false,
                    axis lines=left,
                    ticks=none,
                    xlabel={$\omega$ substituted with equivalent $\mu$},
                    ylabel={cost},
                    domain=0.5:1.75,
                    xmin=0.5, xmax=2,
                    ymin=0, ymax=4,
                    samples=200
                ]
                    \addplot[blue, thick] (x, 2*x^1.1-0.933);
                    \node at (axis cs: 1.925,3) [blue, align=left] {additional\\ system\\ cost};

@@

    \begin{textblock*}{10cm}
        (1.25cm, 1.75cm) % {block width} (coords)
        \centering
        \begin{tikzpicture}
            \draw (4.5,0.0) node[] {\small Opportunity Cost of Wind Power};
        \end{tikzpicture}
    \end{textblock*}

    \begin{textblock*}{10cm}
        (0.75cm, 2.25cm) % {block width} (coords)
        \includegraphics[width=\textwidth, keepaspectratio]{./figures/undisturbed_base}
    \end{textblock*}

    \begin{textblock*}{4.5cm}
        (10.9cm, 2.5cm)
        \begin{small}
            \begin{table}
                \begin{tabular}{p{1.8cm} | r l}
                    Capital cost of solar PV & $630$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}}
                \end{tabular}
            \end{table}
            \smallskip
            \quad $\rightarrow$ about $\sfrac{2}{3}$ rooftop PV, \\
            \quad $\sfrac{1}{3}$ open space PV

            \vspace{3.75cm}
            \quad wind power and solar PV \\ \quad are substitutes
\end{small}
\end{textblock*}

    \begin{textblock*}{9.5cm}
        (1.5cm, 8.7cm)
        \begin{tikzpicture}%[thick]
            \draw (5.75,0) -- (9,0);
            \draw[{Latex[length=2mm]}-] (0,0) -- (3.25,0);
            \draw (4.5,0.0) node[] {\scriptsize more wind power};
        \end{tikzpicture}

@@

\section{Sensitivity}
\begin{frame}
    \frametitle{The cost of undisturbed landscapes}
    \framesubtitle{Sensitivity to capital cost of solar PV}

    \begin{textblock*}{10cm}
        (1.25cm, 1.75cm) % {block width} (coords)
        \centering
        \begin{tikzpicture}
            \draw (4.5,0.0) node[] {\small Least Cost Technology Mix};
        \end{tikzpicture}
    \end{textblock*}

    \begin{textblock*}{10cm}
        (0.75cm, 2.25cm) % {block width} (coords)
%\begin{textblock*}{10cm}(0.75cm, 2.0cm) % {block width} (coords)
        \includegraphics[width=\textwidth, keepaspectratio]{./figures/sens_nrg}
    \end{textblock*}

    \begin{textblock*}{4cm}
        (11cm, 2.5cm)
        \begin{small}
            Capital cost estimates for solar PV in 2030
            \begin{table}
                \begin{tabular}{p{1.8cm} | p{0.3cm} l}
                    Small-scale rooftop & $830$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
                    Utility-scale open space & $280$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
                \end{tabular}
            \end{table}
        \end{small}
    \end{textblock*}
\end{frame}

\begin{frame}
    \frametitle{The cost of undisturbed landscapes}
    \framesubtitle{Sensitivity to capital cost of solar PV}
    \begin{textblock*}{10cm}
        (1.25cm, 1.75cm) % {block width} (coords)
        \centering
        \begin{tikzpicture}
            \draw (4.5,0.0) node[] {\small Opportunity Cost of Wind Power};
        \end{tikzpicture}
    \end{textblock*}

    \begin{textblock*}{10cm}
        (0.75cm, 2.25cm) % {block width} (coords)
%\begin{textblock*}{10cm}(0.75cm, 2.0cm) % {block width} (coords)
        \includegraphics[width=\textwidth, keepaspectratio]{./figures/undisturbed_low}
    \end{textblock*}

    \begin{textblock*}{4cm}
        (11cm, 2.25cm)
        \begin{small}
            \begin{table}
                \begin{tabular}{p{1.8cm} | p{0.3cm} l}
                    Capital cost of solar PV & $280$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}}
                \end{tabular}
            \end{table}
\end{small}
\end{textblock*}
\end{frame}
